---
title: 'Data 607 - Project #1'
author: "Zach Alexander"
date: "9/22/2019"
output: html_document
---

#### Libraries used

```{r, warning=FALSE, comment=FALSE, results=FALSE}
library(stringr)
library(knitr)
```

***

#### Reading the text file from GitHub into R

**First, I read in the text file from my GitHub account - I saved it from Blackboard initially, and then used R to read it into my project code.**

```{r}
chess_data <- paste(readLines("https://raw.githubusercontent.com/zachalexander/data607_cunysps/master/Project1/chess_ratings.txt"), collapse = '\n')
```

***

#### Using Regular Expressions

**Next, I wanted to take the long string and use regular expressions to extract the important information. I initally looked for isolating the player names. Since all of them are in uppercase format, I utilized this as a way to save them into a vector.**

```{r pressure}
player_names <- unlist(str_extract_all(chess_data, '([:upper:]+\\s[:upper:]+(\\s[:upper:]+)?(\\s[:upper:]+)?)'))

# remove the top value since it's not a name
player_names <- player_names[player_names != "USCF ID"]

# change to 'proper' case
player_names <- str_to_title(player_names)
head(player_names)
```

**For Player's State (location), Pre-rating, Total Points, and Post-rating, I also used regular expressions to store this specific information in vectors. See below for my code:**

```{r}
# found that each player's state could be extracted through this regular expression combination
location <- unlist(str_extract_all(chess_data, '(\n)\\s\\s\\s[:upper:]+'))
location <- unlist(str_replace_all(location, '^\\s+|\\s+$', ""))
head(location)
```

```{r}
# found that each player's pre-rating could be extracted through this regular expression combination
pre_rating <- unlist(str_extract_all(chess_data, '(R:)\\s{1,}\\d+'))
pre_rating <- unlist(str_replace_all(pre_rating, '(R:)\\s', ""))
head(pre_rating)
```

```{r}
# found that each player's post-rating could be extracted through this regular expression combination (used later in data visualization)
post_rating <- unlist(str_extract_all(chess_data, '(->)(\\s{1,})?\\d+'))
post_rating <- unlist(str_replace_all(post_rating, '(->)(\\s)?', ""))
```

```{r}
# found that each player's total points could be extracted through this regular expression combination
total_points <- unlist(str_extract_all(chess_data, '(\\d\\.\\d)'))
head(total_points)
```

***

#### Dealing with the player opponent data

**Once the above information was isolated and stored in vectors, I needed to find a way to grab the results from the tournament for each player in a string. To do this, I used a specific regular expression that was able to find the structure of just the result and the opponent number for each players' games. See below: **

```{r}
results <- unlist(str_extract_all(chess_data, '(\\d\\.\\d\\s{2}(\\|[:upper:]\\s{2}(\\s)?(\\s)?(\\d+)?){1,})'))
results <- unlist(str_replace(results, '\\d\\.\\d\\s{2}\\|', ''))
kable(head(results))
```


#### Storing these in a separate data frame

**After grabbing the results, I needed to clean them up a bit in order to eventually replace the opponent number with their pre-tournament elo rating. I then stored this information in a temporary dataframe in order to bind it together with the full data later (stored in 'tempdf'):**
```{r}
opponents <- unlist(str_replace_all(results, '[:upper:]\\s+', ''))
opponents <- unlist(str_replace_all(opponents, '\\|', '\\,'))
opponents <- unlist(str_replace_all(opponents, '\\,{2,}', '\\,'))
opponents <- unlist(str_replace_all(opponents, '(\\,$)', ''))
opponents <- unlist(str_replace_all(opponents, '^\\,', ''))
tempdf <- data.frame(V1 = opponents)
kable(head(tempdf))
```


***

**Once the temporary dataframe had just the opponents data in one column, I wanted to split the strings and create separate columns for each opponent based on the game. To do this, I created a for loop:**

```{r}
# used a for loop here to iterate over V1, using strsplit and the comma to separate out the opponent numbers into the appropriate columns. This also stored N/As in columns where a player didn't play an opponent.
for(i in 1:7){
  tempdf[, paste0('game', i)] <- sapply(strsplit(as.character(tempdf$V1),','), "[", i)
}
kable(head(tempdf))
```


***

#### Merging data into one data frame

**Now that the opponent data was stored correctly in a separate dataframe, I decided to merge this data in with the rest of my vectors to create one large dataframe ('chess_ratings_df'):**

```{r}
chess_ratings_df <- data.frame(player_name = player_names, 
                               player_state = location,
                               total_points = total_points,
                               player_pre_rating = pre_rating, 
                               player_post_rating = post_rating,
                               game1 = tempdf$game1,
                               game2 = tempdf$game2,
                               game3 = tempdf$game3,
                               game4 = tempdf$game4,
                               game5 = tempdf$game5,
                               game6 = tempdf$game6,
                               game7 = tempdf$game7)

kable(head(chess_ratings_df))
```


***


#### Changing the data types from strings to numbers

**In order to do proper average calculations later on, I needed to make sure my data types were correct. For this, I changed the character vectors to numeric vectors:**

```{r}

# just to be safe, I changed the data types for these columns to numbers
chess_ratings_df$player_pre_rating <- as.numeric(as.character(chess_ratings_df$player_pre_rating))
chess_ratings_df$player_post_rating <- as.numeric(as.character(chess_ratings_df$player_post_rating))
chess_ratings_df$total_points <- as.numeric(as.character(chess_ratings_df$total_points))
chess_ratings_df$game1 <- as.numeric(as.character(chess_ratings_df$game1))
chess_ratings_df$game2 <- as.numeric(as.character(chess_ratings_df$game2))
chess_ratings_df$game3 <- as.numeric(as.character(chess_ratings_df$game3))
chess_ratings_df$game4 <- as.numeric(as.character(chess_ratings_df$game4))
chess_ratings_df$game5 <- as.numeric(as.character(chess_ratings_df$game5))
chess_ratings_df$game6 <- as.numeric(as.character(chess_ratings_df$game6))
chess_ratings_df$game7 <- as.numeric(as.character(chess_ratings_df$game7))
```

***

#### Using loops to substitute opponent numbers with their pre-tournament elo ratings

**To substitute the opponent number with their pre-tournament elo rating, I created another set of for loops:**
```{r}
# while iterating over the game columns, iterate over the player opponent number and match it with the opponent's pre-tournament elo rating. Then replace the opponent number with their elo rating.
for(i in 6:12) {
  for(j in 1:64) {
    value <- chess_ratings_df[,i][j]
    chess_ratings_df[,i][j] <- chess_ratings_df$player_pre_rating[value]
  }
}

kable(head(chess_ratings_df))
```


***

#### Taking the average of the pre-tournament elo ratings

**Finally, after the pre-tournament elo ratings had been substituted in, I could then take the average of these ratings across the seven games of the tournament to obtain the Average Pre Chess Rating of Opponents:**

```{r}
chess_ratings_df$average_opp_rating <- round(rowMeans(chess_ratings_df[,6:12], na.rm = TRUE), digits = 0)
```


***

#### For future data visualization work (see at the end), I created a few more columns in the 'chess_ratings_df' data frame.

**For my future data visualization, I decided to create a few more columns. I wanted to see the difference in player pre-tournament rating and player post-tournament rating to see who had the highest elo rating gain. Additionally, I created a column on the conditional of whether or not the rating difference between pre and post was negative or positive. Finally, for charting purposes, I created one final column so that there weren't any negative numbers for the rating differences, by squaring and taking the square root of the difference (will be used later):**
```{r}
chess_ratings_df$rating_difference <- chess_ratings_df$player_post_rating - chess_ratings_df$player_pre_rating
chess_ratings_df$change_pos <- ifelse(chess_ratings_df$rating_difference >=0, TRUE, FALSE)
chess_ratings_df$rating_difference_sqr <- sqrt((chess_ratings_df$rating_difference ^ 2))
```

***
***

#### The final data frame to be exported to a csv file

**In the end, you can see the data frame required for the project ('final_df').**

```{r}
final_df <- chess_ratings_df[, c(1:4, 13)]
kable(final_df)
```


**Here's the syntax for exporting it to my local computer. However, feel free to change the filepath to export it anywhere else.**

```{r}
filename <- '~/Desktop/grad_school/data_607_cunysps/data607_cunysps/Project1/chess_ratings.csv'
write.csv(final_df, file = filename)
```

***
***

#### Data Visualization

**To go a step further with this project, I created a micro-site using a Javascript Framework (Angular), as well as Google Firebase to deploy this to the internet. At this link:** 

**[https://data606-chess.firebaseapp.com/](https://data606-chess.firebaseapp.com/)**  

**...you'll be able to find a 'mock' newspaper article that display a data visualizations and tables using d3.js (charting library for Javascript). The data visualization displays data on hover, as well as more data if you click the associated button -- to show +/- elo ratings.**  

**As a quick note, the url says 'data606' but it should be data607 (I made a typo on the setup - sorry about that!)**

* Steps I used to create my data visualization  
  + Created an Angular project using the angular-cli
  + Created a firebase project and deployed my Angular project using firebase
  + Took my csv file that I created in R, and converted the data to json format
  + Using Javascript and Typescript, I created a d3.js bar chart displaying some of this data
  + I used boostrap (css library) to generate the tables

**If interested, you can find the files for this on my Github as well, at [this repository](https://github.com/zachalexander/data607_chess_elo).**




