---
title: "Homework1"
author: "Zach Alexander"
date: "August 30, 2019"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Loading libraries


```{r, warning=FALSE, message=FALSE}
library(data.table)
library(dplyr)
library(stringr)
library(readr)
```


# Reading the data


**I used read.table to load the data into R from my github repository. I stored the data in a data frame called mushroomData. I used the sep = "," to separate the data by commons found in the raw file.**


```{r}
# pulling the data directly from the source website (commenting out since assignment suggests to pull
# directly from github repository)

# mushroomData <- fread('https://archive.ics.uci.edu/ml/machine-learning-databases/mushroom/agaricus-lepiota.data', header = FALSE, stringsAsFactors = FALSE) #

# pulling data table from my github repository

mushroomData <- "https://raw.githubusercontent.com/zachalexander/data607_cunysps/master/Homework1/agaricus-lepiota.data";

mushroomData <- read.table(mushroomData, sep = ",", stringsAsFactors = FALSE, header = FALSE)

```


# Dimensions of original data and locating edible/poisonous column


**Once I read the data into R, I took a look at its dimensions:**


```{r}
dim(mushroomData)
```


**I also noticed that the first column had values of "e" or "p", which I believed to
characterize mushrooms as either "edible" or "poisonous".**


```{r}
table(mushroomData$V1)
```


**After looking through the documentation, I located the ["Attribute Information" section](https://archive.ics.uci.edu/ml/datasets/Mushroom).
Below, are the column names associated with the dataset. With the prior knowledge that the first column
of the dataset classifies mushrooms as either edible or poisonous, I was able to add in the additional column names after this first column. I decided to store the headers in a vector for later use.**


# Created a vector to store all header information


```{r}
headers <- c('edib-or-poison', 'cap-shape', 'cap-surface', 'cap-color', 'bruises?', 'odor',
             'gill-attachment', 'gill-spacing', 'gill-size', 'gill-color', 'stalk-shape',
             'stalk-root', 'stalk-surface-above-ring', 'stalk-surface-below-ring',
             'stalk-color-above-ring', 'stalk-color-below-ring', 'veil-type', 'veil-color',
             'ring-number', 'ring-type', 'spore-print-color', 'population', 'habitat');
```



# Create data frame to hold all relabeling information



**Created a data frame to hold all relabling values. This information was also taken directly from
the "Attribute Information" section, referenced above. I created a data frame here to utilize later
in a loop.**



```{r}
relabels <- rbind(c('edib-or-poison', "e", "edible"), 
                  c('edib-or-poison', "p", "poisonous"),
                  c('odor', "a", "almond"),
                  c('odor', "l", "anise"),
                  c('odor', "c", "creosote"),
                  c('odor', "y", "fishy"),
                  c('odor', "f", "foul"),
                  c('odor', "m", "musty"),
                  c('odor', "n", "none"),
                  c('odor', "p", "pungent"),
                  c('odor', "s", "spicy"),
                  c('cap-color', "n", "brown"),
                  c('cap-color', "b", "buff"),
                  c('cap-color', "c", "cinnamon"),
                  c('cap-color', "g", "gray"),
                  c('cap-color', "r", "green"),
                  c('cap-color', "p", "pink"),
                  c('cap-color', "u", "purple"),
                  c('cap-color', "e", "red"),
                  c('cap-color', "w", "white"),
                  c('cap-color', "y", "yellow"),
                  c('population', "a", "abundant"),
                  c('population', "c", "clustered"),
                  c('population', "n", "numerous"),
                  c('population', "s", "scattered"),
                  c('population', "v", "several"),
                  c('population', "y", "solitary"),
                  c('habitat', "g", "grasses"),
                  c('habitat', "l", "leaves"),
                  c('habitat', "m", "meadows"),
                  c('habitat', "p", "paths"),
                  c('habitat', "u", "urban"),
                  c('habitat', "w", "waste"),
                  c('habitat', "d", "woods")
                );

relabels <- data.frame(relabels, stringsAsFactors = FALSE)
```



# Added headers to mushroom data



```{r}
for(i in 1:length(headers)) {
  names(mushroomData)[i] <- headers[i]
}
```



# Subset mushroom dataset to include 5 columns


**I picked these columns for the subset because I found that odor can be very predictive as to whether or not a mushroom is ultimately edible or poisonous. Additionally, cap-color has been quite predictive as well. I also thought that adding population information and habitat information may be interesting
for future analysis, given that we could look at trends based on prevelance and types of suitable
habitats.**


```{r}
mushroomData <- select(mushroomData, 'edib-or-poison', 'odor', 'cap-color', 'population', 'habitat')
```


# Utilize the relables data frame to re-label values in mushroom dataset


**By setting up the relabels data frame above, I created a loop to update the data based on its
corresponding values.**

```{r}
for(i in 1:length(relabels$X1)){
  mushroomData[[relabels$X1[i]]] <- replace(mushroomData[[relabels$X1[i]]] , mushroomData[[relabels$X1[i]]] == relabels$X2[i], relabels$X3[i])
}
```



# Print out final, subsetted data frame


```{r}
# just printing the first few rows since it's a long dataset.
head(mushroomData)

# uncomment below to see full dataset 
# mushroomData
```

