---
title: 'DATA 607 - Homework #3 - Zach Alexander'
output: html_document
---

  ***
  
> Libraries used:  
  
```{r, warning=FALSE}
library(stringr)
library(knitr)
```
  
  ***
  
> PROBLEM #3: Copy the introductory example. The vector 'name' stores the extracted names.  

  
  **Reading in variables from the text:**  
```{r}
raw.data <-"555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert"

name <- unlist(str_extract_all(raw.data, "[[:alpha:]., ]{2,}"))
phone <- unlist(str_extract_all(raw.data, "\\(?(\\d{3})?\\)?(-| )?\\d{3}(-| )?\\d{4}"))
```


  ***
  **Problem #3.1. Use the tools of this chapter to rearrange the vector so that all elements conform to the standard first_name last_name.**  

```{r}
## The following steps put the names into first_name last_name order, but also remove second names and titles

# remove the letter indicating the second name ("Burns, C. Mongomery" to "Burns, Montgomery")
name_2 <- unlist(str_replace(name, " \\w{1}\\.", ""))

# for names with commas, locate and flip from last, first to first last
name_3 <- unlist(str_replace(name_2, "(\\w+),\\s(\\w+)", "\\2 \\1"))

# remove titles
just_first_last_name <- unlist(str_replace(name_3, "\\w{1,}\\. ", ""))

# name vector with just first_name and last_name
just_first_last_name
```

```{r}
## The following steps put the names into first_name last_name order, and include second names and titles in proper order

# locate name with a second name and a comma separating last and first, flip the order to first and last ("Burns, C. Montgomery" to "C. Montgomery Burns")
name_5 <- unlist(str_replace(name, "(\\w+),(\\s(\\w).\\s\\w+)", "\\2 \\1"))

# take this same character and flip the first and second parts so the second name is in the middle ("C. Montgomery Burns" to "Montgomery C. Burns")
name_6 <- unlist(str_replace(name_5, "(\\s\\w.\\s)(\\w+)", "\\2\\1"))

# locate name with a comma separating last and first, flip the order to first and last ("Simpson, Homer" to "Homer Simpson")
name_7 <- unlist(str_replace(name_6, "(\\w+),\\s(\\w+)", "\\2 \\1"))

# remove extra spaces between names
first_last_name_extra <- unlist(str_replace(name_7, "\\s{2}", " "))

first_last_name_extra
```


  ***
  **Problem #3.2. Construct a logical vector indicating whether a character has a title (i.e., Rev. and Dr.).**  

```{r}
has_title <- unlist(str_detect(name, "(\\w+).\\s\\w+\\s\\w+"))
has_title
```


  ***
  **Problem #3.3. Construct a logical vector indicating whether a character has a second name.**  
  

```{r}
has_second_name <- unlist(str_detect(name, "\\w+,\\s\\w.\\s\\w+"))
has_second_name
```


  
  **You can find all three solutions in the following data frame:**  
  

```{r}
final_table <- data.frame(name, just_first_last_name, first_last_name_extra, has_title, has_second_name)
kable(final_table)
```

  ***

> PROBLEM #4 Describe the types of strings that conform to the following regular expressions and construct an example that is matched by the regular expression.  


  
  **Problem #4.1**  
```{r}
expression <- "[0-9]+\\$"

# The description of the expression: At least one number followed by a "$" sign.

examples <- c("873212233$", "3$", "abd3ek1$")

# Using str_detect, these should return TRUE if all examples abide by the regular expression criteria
str_detect(examples, expression)
```


  ***
  **Problem #4.2**  
```{r}
expression <- "\\b[a-z]{1,4}\\b"

# The description of the expression: A string of 1 to 4 lowercase letters. The string length has to be at least one lowercase letter, and no more than 4 lowercase letters. There can be no uppercase characters.

examples <- c("dgfg", "road", "help")

# Using str_detect, these should return TRUE if all examples abide by the regular expression criteria
str_detect(examples, expression)
```


  ***
  **Problem #4.3**  
```{r}
expression <- ".*?\\.txt$"

# The description of the expression: Any string pattern, but is also optional, followed by ".txt" at the very end of the string - the "$" denotes the end of the string.

examples <- c(".txt", "dafd$#@;.txt", "adas45322.txt")

# Using str_detect, these should return TRUE if all examples abide by the regular expression criteria
str_detect(examples, expression)
```


  ***
  **Problem #4.4**  
```{r}
expression <- "\\d{2}/\\d{2}/\\d{4}"

# The description of the expression: Two digits, followed by a slash (/), followed by two digits, followed by a slash (/), followed by four digits. Format = xx/xx/xxxx.

examples <- c("12/23/2018", "04/23/1980", "05/12/1990")

# Using str_detect, these should return TRUE if all examples abide by the regular expression criteria
str_detect(examples, expression)
```


  ***  
  **Problem #4.5**  
```{r}
expression <- "<(.+?)>.+?</\\1>"

# The description of the expression: One or more characters inside of <>, followed by another string of characters, followed by a string of characters the same way as the first set, inside of <>, however this set has a "/" at the beginning of the string inside of <>. For instance, this format runs the pattern of html markup <div>Something</div>.

examples <- c("<div>Something</div>", "<section>Test paragraph</section>", "<html><body></body></html>")

# Using str_detect, these should return TRUE if all examples abide by the regular expression criteria
str_detect(examples, expression)
```

  ***
  
  > EXTRA CREDIT: The following code hides a secret message. Crack it with R and regular expressions. Hint: Some of the characters are more revealing that others! The code snippet is also available in the materials at www.r-datacollection.com.
  
  ***
  
###### (I had some help separating the final 'concatString' (see below) by referencing [this source.](https://stackoverflow.com/questions/35542346/r-using-regmatches-to-extract-certain-characters) This helped me identify the periods in the secret code, which I eventually replaced with spaces.)  

  ***
  
```{r}
secret_message <- "clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0TanwoUwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigOd6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr"

# a couple tests using regular expressions

# extract all letters, not case sensitive
all_letters <- unlist(str_extract_all(secret_message, "[[:alpha:]]{1,}"))

# extract all digits
all_digits <- unlist(str_extract_all(secret_message, "[[:digit:]]{1,}"))

# extract all lowercase letters
all_lowercase_letters <- unlist(str_extract_all(secret_message, "[[:lower:]]{1,}"))

# extract all uppercase letters
all_uppercase_letters <- unlist(str_extract_all(secret_message, "[[:upper:]]{1,}"))

# after referring to some help from other sources, I noticed that periods also can be extracted to eventualy remove spaces.
all_uppercase_letters_and_periods <- unlist(str_extract_all(secret_message, "[[:upper:].]"))

# found the secret message! I also noticed that periods can also be extracted to eventually make spaces in the final code. Let's clean it up a bit:
concatString <- paste(all_uppercase_letters_and_periods, collapse = "")

# replace all the periods with spaces to create the final solution:
solution <- str_replace_all(concatString, "[.]", " ")
solution
```
