---
title: "DATA 607 - Week 9 Homework"
author: "Zach Alexander"
date: "10/22/2019"
output: html_document
---

***

#### Libraries used
```{r, warning=FALSE, message=FALSE}
library(httr)
library(knitr)
library(kableExtra)
```

***

#### For this assignment, I used the ["Times Newswire API"](https://developer.nytimes.com/docs/timeswire-product/1/overview)

> This API is set up to hold an up to the minute stream of the latest articles published by the New York Times. 
> This stream holds 20 of the latest articles.


```{r}
# save the url with api-key
url <- "https://api.nytimes.com/svc/news/v3/content/all/all.json?api-key=SoV2EDJdCsYfjVcQeelgFOdkaAlJtPts"
```

***

#### Using the httr library to get the json data from the API
```{r}
nyt <- GET(url)
```


***

#### Parsing the json data, and saving certain parts in vectors
```{r}
details <- content(nyt, "parse")

# creating empty vectors to store results in the following for loop
titles = vector()
abstracts = vector()
sections = vector()
pb_date = vector()
urls = vector()

# iterating through the parsed results and storing data from various sections in vectors to be used later to combine into the final dataframe
for(i in 1:length(details$results)){
  titles[i] <- details$results[[i]]$title
  abstracts[i] <- details$results[[i]]$abstract
  sections[i] <- details$results[[i]]$section
  pb_date[i] <- details$results[[i]]$published_date
  urls[i] <- details$results[[i]]$url
}

```


***

#### Saving to a final data frame

```{r}
# adding all of the vectors to a final dataframe
final_df <- data.frame(titles, abstracts, sections, pb_date, urls)

# printing the final dataframe
kable(final_df, align = rep('c', 5)) %>% 
  kable_styling(bootstrap_options = c("striped"), full_width = F)
```


***

#### Similar work in python

> I also decided to build this out in Python as well, to gain a little more experience working with the Pandas library and python code. If interested, you can find my code at this Jupyter Notebook file on my [GitHub repository](https://github.com/zachalexander/data607_cunysps/tree/master/Homework6) as the .ipynb file.  

> An html version of this file is also at the same location as above.
